---
title: "PS 1 Sample Code"
output:
  html_document:
    df_print: paged
---

```{r}
library(readxl) #reading excel files
library(tidyverse)
library(haven)
library(dplyr)
library(ggplot2)
#install.packages("mFilter")
library(mFilter) #Package used for hp filter
library(cowplot) #Allows for side by side graphs
```

## Read data from Excel file and select relevant fields
```{r}
#Kiersten's WD==getwd()
#install.packages("readxl")
library(readxl)
original <- read_excel("/Users/Kiersten/Documents/School/Session 3/Macro/Measuring_Business_Cycles_HP-Filter/Data/usg_data_annual.xls")
#View(original)
#Select United States and South Korea data
GDP1<- filter(original, Country_Name=="United States" | Country_Name=="Korea, Rep.")

#Filter for relevant business-cycle statistics
GDP2 <- filter(GDP1, Indicator_Code=="NY.GDP.PCAP.KN" | Indicator_Code=="NE.CON.PETC.ZS" | Indicator_Code=="NE.GDI.TOTL.ZS" | Indicator_Code=="NE.CON.GOVT.ZS" | Indicator_Code=="NE.IMP.GNFS.ZS" | Indicator_Code=="NE.EXP.GNFS.ZS")
```

**Restructuring data so year is its own unique variable**
South Korea data
```{r}
#Separating DFs by country to make restructuring simpler (maybe)
SKorea<- filter(GDP2, Country_Name=="Korea, Rep.")

#Remove first three columns in order to transpose
SKorea2 <- t(SKorea)
SKorea3 <- SKorea2[-c(1,2,3),]

#Change values to numeric
SKorea3 <- as.data.frame(SKorea3)

#Rename variables
SKorea4 <- SKorea3 %>%
  select(V1, V2, V3, V4, V5, V6) %>%
  rename(NY.GDP.PCAP.KN = V1, NE.GDI.TOTL.ZS = V2, NE.CON.GOVT.ZS = V3, NE.IMP.GNFS.ZS = V4, NE.EXP.GNFS.ZS = V5, NE.CON.PETC.ZS = V6)

SKorea5 <- SKorea4[-c(1),]

#Add Country variable
Country <- as.list(rep("S. Korea", 52))
length(Country)
country <- as.data.frame(Country)
country2 <- t(country)
SKorea6 <- cbind(SKorea5, country2)

#Name year variable 
SKorea_final <- tibble::rownames_to_column(SKorea6, "YEAR")
SKorea_final <- rename(SKorea_final, COUNTRY = country2)
SKorea_final <- select(SKorea_final, "YEAR", "COUNTRY", "NY.GDP.PCAP.KN", "NE.GDI.TOTL.ZS", "NE.CON.GOVT.ZS", "NE.IMP.GNFS.ZS", "NE.EXP.GNFS.ZS", "NE.CON.PETC.ZS")

#Re-ordering variables and setting to numeric
SKorea_data <- rename(SKorea_final, GDPCapita = NY.GDP.PCAP.KN, HHConPct = NE.CON.PETC.ZS, GCapFormPct = NE.GDI.TOTL.ZS, GovtConPct = NE.CON.GOVT.ZS, ImpPct = NE.IMP.GNFS.ZS, ExpPct = NE.EXP.GNFS.ZS)
SKorea_data$GDPCapita <- as.numeric(SKorea_data$GDPCapita)
SKorea_data$HHConPct <- as.numeric(SKorea_data$HHConPct)
SKorea_data$GCapFormPct <- as.numeric(SKorea_data$GCapFormPct)
SKorea_data$GovtConPct <- as.numeric(SKorea_data$GovtConPct)
SKorea_data$ImpPct <- as.numeric(SKorea_data$ImpPct)
SKorea_data$ExpPct <- as.numeric(SKorea_data$ExpPct)
SKorea_data$YEAR <- as.numeric(SKorea_data$YEAR)
SKorea_data$logGDP<-log(SKorea_data$GDPCapita)

```

US data
```{r}
US<- filter(GDP2, Country_Name=="United States")
#View(US)

#Remove first three rows in order to transpose
US2 <- t(US)
US3 <- US2[-c(1,2,3),]

#Change values to numeric
US3 <- as.data.frame(US3)

#Rename variables
US4 <- US3 %>%
  select(V1, V2, V3, V4, V5, V6) %>%
  rename(NY.GDP.PCAP.KN = V1, NE.GDI.TOTL.ZS = V2, NE.CON.GOVT.ZS = V3, NE.IMP.GNFS.ZS = V4, NE.EXP.GNFS.ZS = V5, NE.CON.PETC.ZS = V6)
US5 <- US4[-c(1),]

#Add Country variable
Countryb <- as.list(rep("US", 52))
length(Countryb)
countryb <- as.data.frame(Countryb)
countryb2 <- t(countryb)
US6 <- cbind(US5, countryb2)

#Name year variable 
US_final <- tibble::rownames_to_column(US6, "YEAR")
US_final <- rename(US_final, COUNTRY = countryb2)
US_final <- select(US_final, "YEAR", "COUNTRY", "NY.GDP.PCAP.KN", "NE.GDI.TOTL.ZS", "NE.CON.GOVT.ZS", "NE.IMP.GNFS.ZS", "NE.EXP.GNFS.ZS", "NE.CON.PETC.ZS")
US_final <- filter(US_final, YEAR >= "1965")

#Re-ordering variables and setting to numeric
US_data <- rename(US_final, GDPCapita = NY.GDP.PCAP.KN, HHConPct = NE.CON.PETC.ZS, GCapFormPct = NE.GDI.TOTL.ZS, GovtConPct = NE.CON.GOVT.ZS, ImpPct = NE.IMP.GNFS.ZS, ExpPct = NE.EXP.GNFS.ZS)
US_data$GDPCapita <- as.numeric(US_data$GDPCapita)
US_data$HHConPct <- as.numeric(US_data$HHConPct)
US_data$GCapFormPct <- as.numeric(US_data$GCapFormPct)
US_data$GovtConPct <- as.numeric(US_data$GovtConPct)
US_data$ImpPct <- as.numeric(US_data$ImpPct)
US_data$ExpPct <- as.numeric(US_data$ExpPct)
US_data$YEAR <- as.numeric(US_data$YEAR)
US_data$logGDP<-log(US_data$GDPCapita)

```

##Merge data sets
```{r}
GDP_data <- rbind(SKorea_data, US_data)
View(GDP_data)
```


**US Business Cycle**
## Use Hodrick-Prescott filter to estimate cylce and secular components of time series.
- De-trend with HP Filter
- Express in terms of percentage deviation from trend
```{r}
#Logged GDP/capita with lambda 100
gdp.hp_US_100<-hpfilter(US_data$logGDP, freq = 100,type = "lambda")
US_data$gdp_T_100<-gdp.hp_US_100$trend
US_data$gdp_C_100<-gdp.hp_US_100$cycle

#Logged GDP/capita with lambda 6.25
gdp.hp_US_6<-hpfilter(US_data$logGDP, freq = 6.25,type = "lambda")
US_data$gdp_T_6<-gdp.hp_US_6$trend
US_data$gdp_C_6<-gdp.hp_US_6$cycle

#Household consumption as pct of GDP with lambda 100
HHCons.hp_US<-hpfilter(US_data$HHConPct, freq = 100,type = "lambda") 
US_data$HHCons_T_100<-HHCons.hp_US$trend
US_data$HHCons_C_100<-HHCons.hp_US$cycle

#Household consumption as pct of GDP with lambda 6.25
HHCons.hp_US<-hpfilter(US_data$HHConPct, freq = 6.25,type = "lambda")
US_data$HHCons_T_6<-HHCons.hp_US$trend
US_data$HHCons_C_6<-HHCons.hp_US$cycle

#Gross capital formation as pct of GDP with lambda 100
GCapForm.hp_US<-hpfilter(US_data$GCapFormPct, freq = 100,type = "lambda") 
US_data$GCapForm_T_100<-GCapForm.hp_US$trend
US_data$GCapForm_C_100<-GCapForm.hp_US$cycle

#Gross capital formation as pct of GDP with lambda 6.25
GCapForm.hp_US<-hpfilter(US_data$GCapFormPct, freq = 6.25,type = "lambda")
US_data$GCapForm_T_6<-GCapForm.hp_US$trend
US_data$GCapForm_C_6<-GCapForm.hp_US$cycle

#Government consumption as pct of GDP with lambda 100
GovtConPct.hp_US<-hpfilter(US_data$GovtConPct, freq = 100,type = "lambda")
US_data$GovtConPct_T_100<-GovtConPct.hp_US$trend
US_data$GovtConPct_C_100<-GovtConPct.hp_US$cycle

#Government consumption as pct of GDP with lambda 6.25
GovtConPct.hp_US<-hpfilter(US_data$GovtConPct, freq = 6.25,type = "lambda")
US_data$GovtConPct_T_6<-GovtConPct.hp_US$trend
US_data$GovtConPct_C_6<-GovtConPct.hp_US$cycle

#Imports as pct of GDP with lambda 100
ImpPct.hp_US<-hpfilter(US_data$ImpPct, freq = 100,type = "lambda")
US_data$ImpPct_T_100<-ImpPct.hp_US$trend
US_data$ImpPct_C_100<-ImpPct.hp_US$cycle

#Imports as pct of GDP with lambda 6.25
ImpPct.hp_US<-hpfilter(US_data$ImpPct, freq = 6.25,type = "lambda")
US_data$ImpPct_T_6<-ImpPct.hp_US$trend
US_data$ImpPct_C_6<-ImpPct.hp_US$cycle

#Exports as pct of GDP with lambda 100
ExpPct.hp_US<-hpfilter(US_data$ExpPct, freq = 100,type = "lambda")
US_data$ExpPct_T_100<-ExpPct.hp_US$trend
US_data$ExpPct_C_100<-ExpPct.hp_US$cycle

#Exports as pct of GDP with lambda 6.25
ExpPct.hp_US<-hpfilter(US_data$ExpPct, freq = 6.25,type = "lambda")
US_data$ExpPct_T_6<-ExpPct.hp_US$trend
US_data$ExpPct_C_6<-ExpPct.hp_US$cycle

View(US_data)
```


**South Korea Business Cycle**
- De-trend with HP Filter
- Express in terms of percentage deviation from trend

```{r}
#Logged GDP/capita with lambda 100
gdp.hp_SK_100<-hpfilter(SKorea_data$logGDP, freq = 100,type = "lambda")
SKorea_data$gdp_T_100<-gdp.hp_SK_100$trend
SKorea_data$gdp_C_100<-gdp.hp_SK_100$cycle

#Logged GDP/capita with lambda 6.25
gdp.hp_SK_6<-hpfilter(SKorea_data$logGDP, freq = 6.25,type = "lambda")
SKorea_data$gdp_T_6<-gdp.hp_SK_6$trend
SKorea_data$gdp_C_6<-gdp.hp_SK_6$cycle

#Household consumption as pct of GDP with lambda 100
HHCons.hp_SK<-hpfilter(SKorea_data$HHConPct, freq = 100,type = "lambda") #not working since US missing data from '60-65
SKorea_data$HHCons_T_100<-HHCons.hp_SK$trend
SKorea_data$HHCons_C_100<-HHCons.hp_SK$cycle

#Household consumption as pct of GDP with lambda 6.25
HHCons.hp_SK<-hpfilter(SKorea_data$HHConPct, freq = 6.25,type = "lambda") #not working since US missing data from '60-65
SKorea_data$HHCons_T_6<-HHCons.hp_SK$trend
SKorea_data$HHCons_C_6<-HHCons.hp_SK$cycle

#Gross capital formation as pct of GDP with lambda 100
GCapForm.hp_SK<-hpfilter(SKorea_data$GCapFormPct, freq = 100,type = "lambda") #not working since US missing data from '60-65
SKorea_data$GCapForm_T_100<-GCapForm.hp_SK$trend
SKorea_data$GCapForm_C_100<-GCapForm.hp_SK$cycle

#Gross capital formation as pct of GDP with lambda 6.25
GCapForm.hp_SK<-hpfilter(SKorea_data$GCapFormPct, freq = 6.25,type = "lambda") #not working since US missing data from '60-65
SKorea_data$GCapForm_T_6<-GCapForm.hp_SK$trend
SKorea_data$GCapForm_C_6<-GCapForm.hp_SK$cycle

#Government consumption as pct of GDP with lambda 100
GovtConPct.hp_SK<-hpfilter(SKorea_data$GovtConPct, freq = 100,type = "lambda")
SKorea_data$GovtConPct_T_100<-GovtConPct.hp_SK$trend
SKorea_data$GovtConPct_C_100<-GovtConPct.hp_SK$cycle

#Government consumption as pct of GDP with lambda 6.25
GovtConPct.hp_SK<-hpfilter(SKorea_data$GovtConPct, freq = 6.25,type = "lambda")
SKorea_data$GovtConPct_T_6<-GovtConPct.hp_SK$trend
SKorea_data$GovtConPct_C_6<-GovtConPct.hp_SK$cycle

#Imports as pct of GDP with lambda 100
ImpPct.hp_SK<-hpfilter(SKorea_data$ImpPct, freq = 100,type = "lambda")
SKorea_data$ImpPct_T_100<-ImpPct.hp_SK$trend
SKorea_data$ImpPct_C_100<-ImpPct.hp_SK$cycle

#Imports as pct of GDP with lambda 6.25
ImpPct.hp_SK<-hpfilter(SKorea_data$ImpPct, freq = 6.25,type = "lambda")
SKorea_data$ImpPct_T_6<-ImpPct.hp_SK$trend
SKorea_data$ImpPct_C_6<-ImpPct.hp_SK$cycle

#Exports as pct of GDP with lambda 100
ExpPct.hp_SK<-hpfilter(SKorea_data$ExpPct, freq = 100,type = "lambda")
SKorea_data$ExpPct_T_100<-ExpPct.hp_SK$trend
SKorea_data$ExpPct_C_100<-ExpPct.hp_SK$cycle

#Exports as pct of GDP with lambda 6.25
ExpPct.hp_SK<-hpfilter(SKorea_data$ExpPct, freq = 6.25,type = "lambda")
SKorea_data$ExpPct_T_6<-ExpPct.hp_SK$trend
SKorea_data$ExpPct_C_6<-ExpPct.hp_SK$cycle

View(SKorea_data)
```

**Business Cycle Statistics**
- US: Calculate the standard deviations, correlation with GDP, and serial correlations (as a percent deviation from the trend)
GDP
```{r}
US_table_100 <- data.frame(matrix(ncol = 6, nrow = 1))
US_table_6 <- data.frame(matrix(ncol = 6, nrow = 1))
GDP_US_100 <- US_data$gdp_C_100*100
GDP_US_6 <- US_data$gdp_C_6*100

US_table_100$GDP_100_sd <- sd(GDP_US_100)
US_table_6$GDP_6_sd <- sd(GDP_US_6)

US_table_100$GDP_100_cor <- cor(GDP_US_100, y=GDP_US_100)
US_table_6$GDP_6_cor <- cor(GDP_US_6, y=GDP_US_6)

#US_table_100$GDP_100_acf <- mean(acf(GDP_US_100, lag=length(GDP_US_100)-1, correlation=TRUE, pl=FALSE))
#US_table_6$GDP_6_acf <- mean(acf(GDP_US_6, lag=length(GDP_US_6)-1, correlation=TRUE, pl=FALSE))
```

Consumption
```{r}
Cons_US_100 <- US_data$HHCons_C_100
Cons_US_6 <- US_data$HHCons_C_6

US_table_100$Cons_100_sd <- sd(Cons_US_100)
US_table_6$Cons_6_sd <- sd(Cons_US_6)

US_table_100$Cons_100_cor <- cor(Cons_US_100, y=GDP_US_100)
US_table_6$Cons_6_cor <- cor(Cons_US_6, y=GDP_US_6)

#ACF_cons_100_US <- as.numeric(acf(Cons_US_100, lag.max=1, correlation=TRUE, pl=FALSE))
#US_table_100$Cons_100_acf <- mean(ACF_cons_100_US)
#ACF_cons_6_US <- as.numeric(acf(Cons_US_100, lag.max=1, correlation=TRUE, pl=FALSE))
#US_table_6$Cons_6_acf <- mean(ACF_cons_100_US)
```

Government spending
```{r}
Govt_US_100 <- ((US_data$GovtConPct_C_100)-(US_data$GovtConPct_T_100))/(US_data$GovtConPct_T_100)
Govt_US_6 <- ((US_data$GovtConPct_C_6)-(US_data$GovtConPct_T_6))/(US_data$GovtConPct_T_6)

US_table_100$Govt_100_sd <- sd(Govt_US_100)
US_table_6$Govt_6_sd <- sd(Govt_US_6)

US_table_100$Govt_100_cor <- cor(Govt_US_100, y=GDP_US_100)
US_table_6$Govt_6_cor <- cor(Govt_US_6, y=GDP_US_6)

#US_table_100$Govt_100_acf <- acf(Govt_US_100, lag=length(Govt_US_100)-1, correlation=TRUE, pl=FALSE)
#US_table_6$Govt_6_acf <- acf(Govt_US_6, lag=length(Govt_US_6)-1, correlation=TRUE, pl=FALSE)
```

Investment
```{r}
Inv_US_100 <- ((US_data$GCapForm_C_100)-(US_data$GCapForm_T_100))/(US_data$GCapForm_T_100)
Inv_US_6 <- ((US_data$GCapForm_C_6)-(US_data$GCapForm_T_6))/(US_data$GCapForm_T_6)

US_table_100$Inv_100_sd <- sd(Inv_US_100)
US_table_6$Inv_6_sd <- sd(Inv_US_6)

US_table_100$Inv_100_cor <- cor(Inv_US_100, y=GDP_US_100)
US_table_6$Inv_6_cor <- cor(Inv_US_6 , y=GDP_US_6)

#US_table_100$Inv_100_acf <- acf(Inv_US_100, lag=length(Inv_US_100)-1, correlation=TRUE, pl=FALSE)
#US_table_6$Inv_6_acf <- acf(Inv_US_6, lag=length(Inv_US_6)-1, correlation=TRUE, pl=FALSE)
```

Imports
```{r}
Imp_US_100 <- ((US_data$ImpPct_C_100)-(US_data$ImpPct_T_100))/(US_data$ImpPct_T_100)
Imp_US_6 <- ((US_data$ImpPct_C_6)-(US_data$ImpPct_T_6))/(US_data$ImpPct_T_6)

US_table_100$Imp_100_sd <- sd(Imp_US_100)
US_table_6$Imp_6_sd <- sd(Imp_US_6)

US_table_100$Imp_100_cor <- cor(Imp_US_100, y=GDP_US_100)
US_table_6$Imp_6_cor <- cor(Imp_US_6, y=GDP_US_6)

#US_table_100$Inv_100_acf <- acf(Inv_US_100, lag=length(Inv_US_100)-1, correlation=TRUE, pl=FALSE)
#US_table_6$Inv_6_acf <- acf(Inv_US_6, lag=length(Inv_US_6)-1, correlation=TRUE, pl=FALSE)
```

Exports
```{r}
Exp_US_100 <- ((US_data$ExpPct_C_100)-(US_data$ExpPct_T_100))/(US_data$ExpPct_T_100)
Exp_US_6 <- ((US_data$ExpPct_C_6)-(US_data$ExpPct_T_6))/(US_data$ExpPct_T_6)

US_table_100$Exp_100_sd <- sd(Exp_US_100)
US_table_6$Exp_6_sd <- sd(Exp_US_6)

US_table_100$Exp_100_cor <- cor(Exp_US_100, y=GDP_US_100)
US_table_6$Exp_6_cor <- cor(Exp_US_6, y=GDP_US_6)

#US_table_100$Inv_100_acf <- acf(Inv_US_100, lag=length(Inv_US_100)-1, correlation=TRUE, pl=FALSE)
#US_table_6$Inv_6_acf <- acf(Inv_US_6, lag=length(Inv_US_6)-1, correlation=TRUE, pl=FALSE)
```


- South Korea: Calculate the standard deviations, correlation with GDP, and serial correlations (as a percent deviation from the trend)
GDP
```{r}
SK_table_100 <- data.frame(matrix(ncol = 6, nrow = 1))
SK_table_6 <- data.frame(matrix(ncol = 6, nrow = 1))
GDP_SK_100 <- SKorea_data$gdp_C_100*100
GDP_SK_6 <- SKorea_data$gdp_C_6*100

SK_table_100$GDP_100_sd <- sd(GDP_SK_100)
SK_table_6$GDP_6_sd <- sd(GDP_SK_6)

SK_table_100$GDP_SK_100_cor <- cor(GDP_SK_100, y=GDP_SK_100)
SK_table_6$GDP_SK_6_cor <- cor(GDP_SK_6, y=GDP_SK_6)

#SK_table_100$GDP_SK_100_acf <- acf(GDP_SK_100, lag=length(GDP_SK_100)-1, correlation=TRUE, pl=FALSE)
#SK_table_6$GDP_SK_6_acf <- acf(GDP_SK_6, lag=length(GDP_SK_6)-1, correlation=TRUE, pl=FALSE)
```

Consumption
```{r}
Cons_SK_100 <- ((SKorea_data$HHCons_C_100)-(SKorea_data$HHCons_T_100))/(SKorea_data$HHCons_T_100)
Cons_SK_6 <- ((SKorea_data$HHCons_C_6)-(SKorea_data$HHCons_T_6))/(SKorea_data$HHCons_T_6)

SK_table_100$Cons_100_sd <- sd(Cons_SK_100)
SK_table_6$Cons_6_sd <- sd(Cons_SK_6)

SK_table_100$Cons_100_cor <- cor(Cons_SK_100, y=GDP_SK_100)
SK_table_6$Cons_6_cor <- cor(Cons_SK_6, y=GDP_SK_6)

#SK_table_100$Cons_100_acf <- acf(Cons_SK_100, lag=length(Cons_SK_100)-1, correlation=TRUE, pl=FALSE)
#SK_table_6$Cons_6_acf <- acf(Cons_SK_6, lag=length(Cons_SK_6)-1, correlation=TRUE, pl=FALSE)
```

Government spending
```{r}
Govt_SK_100 <- ((SKorea_data$GovtConPct_C_100)-(SKorea_data$GovtConPct_T_100))/(SKorea_data$GovtConPct_T_100)
Govt_SK_6 <- ((SKorea_data$GovtConPct_C_6)-(SKorea_data$GovtConPct_T_6))/(SKorea_data$GovtConPct_T_6)

SK_table_100$Govt_100_sd <- sd(Govt_SK_100)
SK_table_6$Govt_6_sd <- sd(Govt_SK_6)

SK_table_100$Govt_100_cor <- cor(Govt_SK_100, y=GDP_SK_100)
SK_table_6$Govt_6_cor <- cor(Govt_SK_6, y=GDP_SK_6)

#SK_table$Govt_100_acf <- acf(Govt_SK_100, lag=length(Govt_SK_100)-1, correlation=TRUE, pl=FALSE)
#SK_table_6$Govt_6_acf <- acf(Govt_SK_6, lag=length(Govt_SK_6)-1, correlation=TRUE, pl=FALSE)
```

Investment
```{r}
Inv_SK_100 <- ((SKorea_data$GCapForm_C_100)-(SKorea_data$GCapForm_T_100))/(SKorea_data$GCapForm_T_100)
Inv_SK_6 <- ((SKorea_data$GCapForm_C_6)-(SKorea_data$GCapForm_T_6))/(SKorea_data$GCapForm_T_6)

SK_table_100$Inv_100_sd <- sd(Inv_SK_100)
SK_table_6$Inv_6_sd <- sd(Inv_SK_6)

SK_table_100$Inv_100_cor <- cor(Inv_SK_100, y=GDP_SK_100)
SK_table_6$Inv_6_cor <- cor(Inv_SK_6 , y=GDP_SK_6)                                                      

#SK_table$Inv_100_acf <- acf(Inv_SK_100, lag=length(Inv_SK_100)-1, correlation=TRUE, pl=FALSE)
#SK_table_6$Inv_6_acf <- acf(Inv_SK_6, lag=length(Inv_SK_6)-1, correlation=TRUE, pl=FALSE)
```

Imports
```{r}
Imp_SK_100 <- ((SKorea_data$ImpPct_C_100)-(SKorea_data$ImpPct_T_100))/(SKorea_data$ImpPct_T_100)
Imp_SK_6 <- ((SKorea_data$ImpPct_C_6)-(SKorea_data$ImpPct_T_6))/(SKorea_data$ImpPct_T_6)

SK_table_100$Imp_100_sd <- sd(Imp_SK_100)
SK_table_6$Imp_6_sd <- sd(Imp_SK_6)

SK_table_100$Imp_100_cor <- cor(Imp_SK_100, y=GDP_SK_100)
SK_table_6$Imp_6_cor <- cor(Imp_SK_6 , y=GDP_SK_6)

#SK_table_100$Inv_100_acf <- acf(Inv_SK_100, lag=length(Inv_SK_100)-1, correlation=TRUE, pl=FALSE)
#SK_table_6$Inv_6_acf <- acf(Inv_SK_6, lag=length(Inv_SK_6)-1, correlation=TRUE, pl=FALSE)
```

Exports
```{r}
Exp_SK_100 <- ((SKorea_data$ExpPct_C_100)-(SKorea_data$ExpPct_T_100))/(SKorea_data$ExpPct_T_100)
Exp_SK_6 <- ((SKorea_data$ExpPct_C_6)-(SKorea_data$ExpPct_T_6))/(SKorea_data$ExpPct_T_6)

SK_table_100$Exp_100_sd <- sd(Exp_SK_100)
SK_table_6$Exp_6_sd <- sd(Exp_SK_6)

SK_table_100$Exp_100_cor <- cor(Exp_SK_100, y=GDP_SK_100)
SK_table_6$Exp_6_cor <- cor(Exp_SK_6, y=GDP_SK_6)

#SK_table_100$Exp_100_acf <- acf(Exp_SK_100, lag=length(Exp_SK_100)-1, correlation=TRUE, pl=FALSE)
#SSK_table_6$Exp_6_acf <- acf(Exp_SK_6, lag=length(Exp_SK_6)-1, correlation=TRUE, pl=FALSE)
```

**Combine and visualize business cycle statistics**
Combining data
```{r}
#Lambda=100
#Create combined table with US and SK as rows 
Ut1 <- t(US_table_100)
Kt1 <- t(SK_table_100)
SK_table21 <- as.data.frame(Kt1)
US_table21 <- as.data.frame(Ut1)

#Rename variable column
USdatatable1 = US_table21 %>% rename(US = V1)
SKdatatable1 = SK_table21 %>% rename(SK = V1)
SK_table_final_100 <- as.data.frame(SKdatatable1[-c(1:6), ])
US_table_final_100 <- as.data.frame(USdatatable1[-c(1:6), ])
View(US_table_final_100)
View(SK_table_final_100)

#Combine two tables
Table_data100 <- cbind(US_table_final_100, SK_table_final_100)
F_table_100 <- as.data.frame(Table_data100)
Final_table_100 <- tibble::rownames_to_column(F_table_100, "MEASURE")
Final_table_100$lambda <- rep(100, 12)
View(Final_table_100)
```

```{r}
#Lambda=6.25
#Create combined table with US and SK as rows 
Ut6 <- t(US_table_6)
Kt6 <- t(SK_table_6)
SK_table6 <- as.data.frame(Kt6)
US_table6 <- as.data.frame(Ut6)

#Rename variable column
USdatatable6 = US_table6 %>% rename(US = V1)
SKdatatable6 = SK_table6 %>% rename(SK = V1)
SK_table_final6 <- SKdatatable6[-c(1:6), ]
US_table_final6 <- USdatatable6[-c(1:6), ]
#View(SK_table_final6)
#View(US_table_final6)

#Combine two tables
Table_data6 <- cbind(US_table_final6, SK_table_final6)
F_table_6 <- as.data.frame(Table_data6)
Final_table_6 <- tibble::rownames_to_column(F_table_6, "MEASURE")
Final_table_6$lambda <- rep(6, 12)
View(Final_table_6)

#Combine final tables
Final_table6 <- rbind(Final_table_6, Final_table_100)
View(Final_table6)
```


Package for table configuration
```{r}
install.packages("kableExtra")
library(kableExtra)
install.packages("knitr")
library(knitr)
```

Merging into table
```{r}
#Make StatID variable
Final_table$STAT <- rep(c("sd","cor"), 12)
#Ordering rows alphabetically by STAT
Final_table <- Final_table[order(Final_table$STAT), ]
rownames(Final_table)=NULL
View(Final_table)

#Table function
base <- Final_table %>%
    kbl(align=c(rep("c", 7), rep("r", 5)))

base %>%
    kable_styling(bootstrap_options = c("striped", "bordered", "condensed", "hover", "responsive"))
```

```{r}
kable_styling(bootstrapkbl(Final_table_ordered, caption = "Group Rows") %>%
  kable_paper("striped", full_width = F) %>%
  pack_rows("Group 1", 5, 7) %>%
  pack_rows("Group 2", 5, 10) %>%
  pack_rows("Group 2", 5, 10)
```


**Business Cycle Visualizations**
## Draw two figures. 
- On the first, draw the raw data and the secular trend estimate. 
- On the second draw the GDP cycle estimate.

#United States
```{r}
plot1<-ggplot(data=US_data)+ 
  geom_line(aes(x=YEAR,y=logGDP,color='Log GDP', group=1),size=1.5) + 
  geom_line(aes(x=YEAR,y=gdp_T_100,color='Log GDP Secular', group=1),linetype=1,size=1.5) +
  xlab("Year")+
  ylab("")+
  scale_colour_manual(name="",values=c('Log GDP'='blue','Log GDP Secular'='red'))+
  ggtitle('US GDP Trend - Lambda=100')+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
plot2<-ggplot(data=US_data)+ 
geom_line(aes(x=YEAR,y=gdp_C_100,color='Log GDP Cycle'),size=1.5,linetype=2)  +xlab("Year")+ylab("")+scale_colour_manual(name="",values=c('Log GDP Cycle'='blue'))+ggtitle('GDP Cycle')+theme(plot.title = element_text(hjust = 0.5))
```

```{r}
plot3<-ggplot(data=US_data)+ 
  geom_line(aes(x=YEAR,y=logGDP,color='Log GDP', group=1),size=1.5) + 
  geom_line(aes(x=YEAR,y=gdp_T_6,color='Log GDP Secular', group=1),linetype=1,size=1.5) +
  xlab("Year")+
  ylab("")+
  scale_colour_manual(name="",values=c('Log GDP'='blue','Log GDP Secular'='red'))+
  ggtitle('US GDP Trend - Lambda=6.25')+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
plot4<-ggplot(data=US_data)+ 
geom_line(aes(x=YEAR,y=gdp_C_6,color='Log GDP Cycle'),size=1.5,linetype=2)  +xlab("Year")+ylab("")+scale_colour_manual(name="",values=c('Log GDP Cycle'='blue'))+ggtitle('GDP Cycle')+theme(plot.title = element_text(hjust = 0.5))
```


```{r}
plot_grid(plot1, plot2, ncol = 1,align = 'v')
```

```{r}
plot_grid(plot3, plot4, ncol = 1,align = 'v')
```

#South Korea
```{r}
plot5<-ggplot(data=SKorea_data)+ 
  geom_line(aes(x=YEAR,y=logGDP,color='Log GDP', group=1),size=1.5) + 
  geom_line(aes(x=YEAR,y=gdp_T_100,color='Log GDP Secular', group=1),linetype=1,size=1.5) +
  xlab("Year")+
  ylab("")+
  scale_colour_manual(name="",values=c('Log GDP'='blue','Log GDP Secular'='red'))+
  ggtitle('South Korea GDP Trend - Lambda=100')+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
plot6<-ggplot(data=SKorea_data)+ 
geom_line(aes(x=YEAR,y=gdp_C_100,color='Log GDP Cycle'),size=1.5,linetype=2)  +xlab("Year")+ylab("")+scale_colour_manual(name="",values=c('Log GDP Cycle'='blue'))+ggtitle('GDP Cycle')+theme(plot.title = element_text(hjust = 0.5))
```

```{r}
plot7<-ggplot(data=SKorea_data)+ 
  geom_line(aes(x=YEAR,y=logGDP,color='Log GDP', group=1),size=1.5) + 
  geom_line(aes(x=YEAR,y=gdp_T_6,color='Log GDP Secular', group=1),linetype=1,size=1.5) +
  xlab("Year")+
  ylab("")+
  scale_colour_manual(name="",values=c('Log GDP'='blue','Log GDP Secular'='red'))+
  ggtitle('South Korea GDP Trend - Lambda=6.25')+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
plot8<-ggplot(data=SKorea_data)+ 
geom_line(aes(x=YEAR,y=gdp_C_6,color='Log GDP Cycle'),size=1.5,linetype=2)  +xlab("Year")+ylab("")+scale_colour_manual(name="",values=c('Log GDP Cycle'='blue'))+ggtitle('GDP Cycle')+theme(plot.title = element_text(hjust = 0.5))
```

```{r}
plot_grid(plot5, plot6, ncol = 1,align = 'v')
```

```{r}
plot_grid(plot7, plot8, ncol = 1,align = 'v')
```

## Discussion
- Discuss how the detrending method influences the volatility of the cyclical component of output

Using a larger lambda results in smoother business cycles and does not emphasize as much GDP fluctuation, 

- Discuss which detrending method identifies recessions for the United States most in line with the NBER business-cycle dates
